Architectures and Open pNets

### 2018-03-07 2018-Arch-pNets/arch-pNets.tex
Guards and leaves (reu)
Changements durant skype simon
Before skype: shorter def 11 and cleaned up a bit expression (now tex compiles)

We first define the set of actions a pLTS can use, let $a$
range over action labels\removeSB{, $\symb{op}$ are operators, and $x_i$ range over
variable names}. Action terms are\replaceSB{:}{{} of the form $\alpha=a((?x_i)^{i \in I}, \symb{Expr}_j^{j \in J})$, where $(?x_i)^{i \in I}$ are input variables, $\symb{Expr}_j^{j \in J})$ are expressions.}\noteLH{Eric suggests to remove this: too precise. I 
think last Expression should be removed not sure about the rest -- to be discussed}\noteSB{I agree: expressions are not needed, but it is nice to explain how action terms are formed }
\removeSBin{\[
\begin{array}[l]{rcl@{\quad}p{5.5cm}}
  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
  p_i&::=& ?x~|~\symb{Expr}&\text{parameters (input variable or expression)}\\
  \symb{Expr}&::=& 
  \symb{Value}~|~x~|~\symb{op}(\symb{Expr}_1,..,\symb{Expr}_n)&\text{Expressions}
\end{array}
\]}
\removeSB{The input variables in an action term are those marked with a
$\symb{?}$.}
We additionally suppose that each input variable does not
appear \replaceSB{somewhere else }{anywhere }in the same action term\replace[Maybe? Not sure this is necessary.\Simon]{:
$p_i=?x\Rightarrow\forall j\neq i.\, x\notin \vars(p_j)$}{.}

@@

A composite pNet consists of a set of sub-pNets exposing
a set of actions, each of them synchronising actions in each of
the sub-pNets. The synchronisation between global actions and
internal actions is given by  \emph{synchronisation vectors}: a
synchronisation vector synchronises one or several internal actions, and
exposes a single resulting global action.
Actions involved at the pNet level (in the synchronisation vectors) do
not \replaceSB{need to distinguish  input 
variables. Action terms for pNets are defined as follows:}{have input variables, \ie they have the form $a(e_j^{j \in J})$.}
\removeSBin{
\[\begin{array}[l]{rcl@{\quad}l}
  \alpha\in \AlgAS &::=&a(Expr_1,\ldots,Expr_n)
\end{array}
\]
}

@@

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined 
inductively as:
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i=\pLTS\}}
\end{array}\]
\end{itemize}
\end{definition}

A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
is said to be \emph{open}.

@@

\noteLH{a pred is a g a post is a e}
The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where \noteSB{This might give a false impression that there is some kind of CCS-like ``duality'' between pLTSs and holes}\hl{each transition composes transitions of several LTSs with
actions of some holes}, the transition occurs if some predicates hold, and can involve a 
set of state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a set \noteSB{In the architecture section, I say that we omit indices on $\goesto{}$ and that they are always clear from the context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
	I}$} of LTSs, a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
	{s \OTarrow {v}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and for all
        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and $\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort $\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
        is a transition of the hole $j$}, for any action $b_j$ in the
        sort $\Sort_j$.  \noteSB{Why $v$? Why not $\alpha$? As now, it is not clear whether this is an action term or a variable $v \in \vars(s)$ of some hypothetical state $s$.}\hl{$v$ is a variable denoting the resulting action}
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $\{x_k\gets e_k\}^{k\in K}$ 
	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
	variables of the open transition.
\end{definition}

@@

\paragraph{Predicates:}
Let
$\mylangle\overline{\pNet},\overline{\Sort},\overline{\symb{SV}}\myrangle$
be a pNet. Consider a synchronisation vector $\symb{SV}\in \overline{\symb{SV}}$. We 
define a
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\begin{multline}
  \Pred_{\symb{sv}}(SV, a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow
  \\
%%   \bigg(
%% \begin{array}{l}
\exists {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.\, SV=\bigl({(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v' | g\bigr)
\\
\land
\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v' \land g
%% \end{array} \bigg)
\end{multline}
\noteLH{Is and g sufficient? I would say yes}
\noteSB{In principle, yes. I am not 100\% sure that the implicit dependency of $g$ on the variables in $a'_i$ and $b'_j$ is clear.}

@@

	\noindent
    $\inferrule
    {k\!\in\! K \\SV_k \!=\! \alpha_m^{m \in I\uplus J} \!\to\! 
    \alpha' |g \\
    	Leaves(\mylangle \pNet_i^{i\in I}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\!\!\in\!\! I. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
\end{array}\right)}\\
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
     J' = \biguplus_{m\in I}\!\! J'_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
    		I' = \biguplus_{m\in I}\!\! I_m'
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
    }
    {\mylangle \pNet_i^{i\in I}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}

>>>>

We first define the set of actions a pLTS can use, let $a$
range over action labels\removeSB{, $\symb{op}$ are operators, and $x_i$ range over
variable names}. Action terms are\replaceSB{:}{{} of the form $\alpha=a((?x_i)^{i \in I}, 
\Expr_j^{j \in J})$, where $(?x_i)^{i \in I}$ are input variables, $\Expr_j^{j \in J})$ 
are expressions.}\noteLH{Eric suggests to remove this: too precise. I 
think last Expression should be removed not sure about the rest -- to be discussed}\noteSB{I agree: expressions are not needed, but it is nice to explain how action terms are formed }
\removeSBin{def of $\AlgA$, $p_i$, $\Expr$}
%\[
%\begin{array}[l]{rcl@{\quad}p{5.5cm}}
%  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
%  p_i&::=& ?x~|~\Expr&\text{parameters (input variable or expression)}\\
%  \Expr&::=& 
%  \symb{Value}~|~x~|~\symb{op}(\Expr_1,..,\Expr_n)&\text{Expressions}
%\end{array}
%\]
\removeSB{The input variables in an action term are those marked with a
$\symb{?}$.}
We additionally suppose that each input variable does not
appear \replaceSB{somewhere else }{anywhere }in the same action term\replace[Maybe? Not sure this is necessary.\Simon]{:
$p_i=?x\Rightarrow\forall j\neq i.\, x\notin \vars(p_j)$}{.}

@@

A composite pNet consists of a set of sub-pNets exposing
a set of actions, each of them synchronising actions in each of
the sub-pNets. The synchronisation between global actions and
internal actions is given by  \emph{synchronisation vectors}: a
synchronisation vector synchronises one or several internal actions, and
exposes a single resulting global action.
Actions involved at the pNet level (in the synchronisation vectors) do
not \replaceSB{need to distinguish  input 
variables. Action terms for pNets are defined as follows:}{have input variables, \ie they 
have the form $a(\Expr_j^{j \in J})$.}
\removeSBin{
%\[
%\begin{array}[l]{rcl@{\quad}l}
$  \alpha\in \AlgAS ::=a(\Expr_1,\ldots,\Expr_n)$
%\end{array}
%\]
}

@@

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined 
inductively as:
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i\pNet_i \text{ is a }\pLTS\}}
\end{array}\]
\end{itemize}
\end{definition}

\noteLHin{Below a simpler def 11}
\begin{definition}[Sorts, Holes, Leaves of pNets]
  \begin{itemize}
  \item 
 The sort of a pNet is its signature, i.e. the set of actions it can
perform. A sort is a set of action signature, where each action signature is an action 
label plus the arity of the action.

\item
The set of holes $\Holes(\pNet_{i_1})$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets (we suppose those indexes 
disjoints).
%
%defined inductively; the sets of holes
%in a pNet node and its subnets are all disjoint:
%  \[\begin{array}{l}
%\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
%\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
%=J\uplus{\displaystyle \bigcup_{i\in 
%I}\Holes(\pNet_i)}\\
%\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
%\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
%\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
%\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(\pNet)= \mylangle \pNet_i \myrangle^{i \in L}$.
is said to be \emph{open}.

@@

\noteLH{a pred is a g a post is a e}
The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where \noteSB{This might give a false impression that there is some kind of CCS-like ``duality'' between pLTSs and holes}\hl{each transition composes transitions of several LTSs with
actions of some holes}, the transition occurs if some predicates hold, and can involve a 
set of state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a set \noteSB{In the architecture section, I say that we omit indices on $\goesto{}$ and that they are always clear from the context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
	I}$} of LTSs, a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
	{s \OTarrow {v}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and for all
        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and $\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort $\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
        is a transition of the hole $j$}, for any action $b_j$ in the
        sort $\Sort_j$.  \noteSB{Why $v$? Why not $\alpha$? As now, it is not clear whether this is an action term or a variable $v \in \vars(s)$ of some hypothetical state $s$.}\hl{$v$ is a variable denoting the resulting action}
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
	would like to just put $\Post_=e$}
	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
	variables of the open transition.
\end{definition}

@@

\paragraph{Predicates:}
Let
$\mylangle\overline{\pNet},\overline{\Sort},\overline{\symb{SV}}\myrangle$
be a pNet. Consider a synchronisation vector $\symb{SV}\in \overline{\symb{SV}}$. We 
define a
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\begin{multline}
  \Pred_{\symb{sv}}(SV, 
  a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow
  \\
%\exists 
%%   \bigg(
%% \begin{array}{l}
\exists {\vars((a'_i)}^{i\in I}),
{\vars((b'_j)}^{j\in J}),\vars(v').\, %SV=
%\\
%\land
\\\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v' \land g
%% \end{array} \bigg)
\\
\text{ Where } SV=\bigl({(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v' | g\bigr)
\end{multline}
\noteLH{Is and g sufficient? I would say yes}
\noteSB{In principle, yes. I am not 100\% sure that the implicit dependency of $g$ on the variables in $a'_i$ and $b'_j$ is clear.}

@@

	\noindent
    $\inferrule
    {k\!\in\! K \\SV_k \!=\! \alpha_m^{m \in I\uplus J} \!\to\! 
    \alpha' |g \\
    	Leaves(\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\!\!\in\!\! I. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
\end{array}\right)}\\
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
     J' = \biguplus_{m\in I}\!\! J'_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
    		I' = \biguplus_{m\in I}\!\! I_m'
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
    }
    {\mylangle \pNet_i^{i\in I'}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}
